### 2012-11-07 paper_sean/introduction.tex
THIS COMMIT CONTAINS DOUBLE-BLIND REVIEW CHANGES

% !TEX root =main.tex
\figgp{quadratotWhiteBooties}{.45}{quadratot_simulator}{.45}{The QuadraTot robot platform on which gaits were evolved. \textbf{Left: } The physical robot, which is composed of 3-D printable and off-the-shelf components.
\textbf{Right: } The robot in the simulator.}

@@

Previous work has shown that the Hypercube-based NeuroEvolution of Augmenting Topologies (HyperNEAT) generative encoding~\cite{stanley2009hypercube} can automatically generate a variety of regular gaits that outperform gaits evolved with direct encodings~\cite{clune2009evolving,clune2011performance}.
However, that work only verified these claims in simulation. 
Yosinski et al. evolved gaits with HyperNEAT directly in hardware on the QuadraTot robot platform~(\figref{quadratotWhiteBooties}). They found that HyperNEAT's gaits outperformed manually designed, parameterized learning algorithms, but still did not produce impressive, natural gaits\footnote{Videos available at http://creativemachines.cornell.edu/evolved-quadruped-gaits}~\cite{yos:clune}. 
A follow-up study built a simulator for QuadraTot to test whether the inclusion of a simulator would improve results and found that it did:  when gaits were evolved in this simulator with a simple direct encoding and then transferred to the QuadraTot robot, the resulting gaits were faster than the gaits produced by evolving gaits with HyperNEAT directly on the robot~\cite{glette}. 
The simulator helped because it afforded much larger population sizes and more generations than were possible when evolving directly in hardware, resulting in 333 times more evaluations per run (60000 vs. 180)~\cite{yos:clune,glette}.

>>>>

% !TEX root =main.tex
\figgp{quadratotWhiteBooties}{.45}{quadratot_simulator}{.45}{The QuadraTot robot platform on which gaits were evolved. \textbf{Left: } The physical robot, which is composed of 3-D printable and off-the-shelf components.
\textbf{Right: } The representation of the robot in the simulator.}

@@

Previous work has shown that the Hypercube-based NeuroEvolution of Augmenting Topologies (HyperNEAT) generative encoding~\cite{stanley2009hypercube} can automatically generate a variety of regular gaits that outperform gaits evolved with direct encodings~\cite{clune2009evolving,clune2011performance}.
However, that work only verified these claims in simulation. 
Yosinski et al. evolved gaits with HyperNEAT directly in hardware on the QuadraTot robot platform~(\figref{quadratotWhiteBooties}). They found that HyperNEAT's gaits outperformed manually designed, parameterized learning algorithms, but still did not produce impressive, natural gaits~\cite{yos:clune}.\footnote{Videos available at (Link removed for double-blind review)}
A follow-up study built a simulator for QuadraTot to test whether the inclusion of a simulator would improve results and found that it did:  when gaits were evolved in simulation with a simple direct encoding and then transferred to the QuadraTot robot, the resulting gaits were faster than the gaits produced by evolving gaits with HyperNEAT directly on the robot~\cite{glette}. 
The simulator helped because it afforded much larger population sizes and more generations than were possible when evolving directly in hardware, resulting in 333 times more evaluations per run (60000 vs. 180)~\cite{yos:clune,glette}.

### 2012-11-07 paper_sean/results.tex
THIS COMMIT CONTAINS DOUBLE-BLIND REVIEW CHANGES

On the physical robot, HyperNEAT gaits from this study outperformed gaits from all previous QuadraTot studies~(\tabref{results})~\cite{yos:clune,haocheng,glette}, including those of Glette et al. 
 However, comparing performance in hardware between studies performed in
different laboratories is difficult, not only because reality is
inherently noisy, but because even two copies of the same robot are
not identical and may produce different speeds for the same input
gait. Gaits for this study and two previous studies~\cite{yos:clune,haocheng} were evaluated on a copy of the QuadraTot robot in the Cornell Creative Machines Lab (CCML), while the gaits in Glette et al. were evaluated on a different copy of the same robot in the Robotics and
Intelligent Systems (ROBIN) lab at the University of Oslo. The two robots were evaluated on two different surfaces, and had different material enveloping their feet to increase friction (compare \figref{quadratotWhiteBooties}-Left to Figure 2 of Glette et al).

The previous fastest gait on any copy of a QuadraTot was from Glette
et al., and traveled 17.8 cm/s on the ROBIN QuadraTot. The
fastest gait produced by HyperNEAT with a simulator in the experiments for this paper traveled 14.5
cm/s on the CCML QuadraTot. It was unclear whether this difference in performance was due to
the differences in the gaits themselves or dissimilarities between
hardware. To control for this possibility, we ran the fastest gait
from Glette et al. 10 times on CCML and measured an average speed of
only 12.95 cm/s $\pm0.93$ SD (vs. 17.8 cm/s measured on ROBIN) and a maximum of
13.8 cm/s (\tabref{results}). It thus appears that the CCML version of the robot is slower. Moreover, the HyperNEAT gait (14.5 cm/s) is faster than the best gait of Glette et al. on the CCML QuadraTot. Unfortunately, it was not possible to test the best HyperNEAT gait on the ROBIN QuadraTot. Because HyperNEAT outperformed the simple encoding from Glette et al. on the same robotic hardware, we tentatively
conclude that HyperNEAT produces faster gaits for the QuadraTot robot. This conclusion is supported by the fact that HyperNEAT also outperformed the encoding from Glette et al. in simulation.

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline
                                         &              & Simulated  & Real Vel. & Real Vel.  \\
                                         & Evaluations  & Velocity &    (CCML)     & (ROBIN) \\
\hline
Parameterized gaits + optimization \cite{yos:clune}   &153    & --    & 5.8 & --\\
\hline
HyperNEAT in hardware \cite{yos:clune}                 & 180         & --         &   9.7  & --   \\
\hline
RL PoWER Spline \cite{haocheng}                         & 300         & --         &   11.1 & --\\
\hline
GA + simulator \cite{glette}             & 60000       & *16.7       &   13.8   & \textbf{17.8}  \\
\hline
HyperNEAT + simulator [this paper]                     & 40000       & \textbf{25.4}       &   \textbf{14.5} & --\\
\hline
\end{tabular}
\vspace{.35cm}
\caption{Velocities of evolved gaits in simulation and on two different copies of the QuadraTot robot. Subject to availability, data are reported from the experiments for this paper and three previous studies. Reported are the total number of evaluations per run, the average of the fastest gaits produced in each run in simulation, and the single fastest gait produced on the CCML and ROBIN copies of the QuadraTot robot (see text for their differences). *Instead of using the single fitness value reported in~\cite{glette}, we ran 19 additional runs and used the average fitness of those 20 runs. Velocities are in cm/s, and bold indicates the best performance. }  
\tablabel{results}
\end{center}
\end{table}

>>>>

On the physical robot, HyperNEAT gaits from this study outperformed gaits from all previous QuadraTot studies~(\tabref{results})~\cite{yos:clune,haocheng,glette}, including those of Glette et al. 
 However, comparing performance in hardware between studies performed in
different laboratories is difficult, not only because reality is
inherently noisy, but because even two copies of the same robot are
not identical and may produce different speeds for the same input
gait. Gaits for this study and two previous studies~\cite{yos:clune,haocheng} were evaluated on a copy of the QuadraTot robot in the (Name removed for double-blind review) Lab (---L), while the gaits in Glette et al. were evaluated on a different copy of the same robot in the Robotics and
Intelligent Systems (ROBIN) lab at the University of Oslo. The two robots were evaluated on two different surfaces, and had different material enveloping their feet to increase friction (compare \figref{quadratotWhiteBooties}-Left to Figure 2 of Glette et al).

The previous fastest gait on any copy of a QuadraTot was from Glette
et al., and traveled 17.8 cm/s on the ROBIN QuadraTot. The
fastest gait produced by HyperNEAT with a simulator in the experiments for this paper traveled 14.5
cm/s on the ---L QuadraTot. It was unclear whether this difference in performance was due to
the differences in the gaits themselves or dissimilarities between
hardware. To control for this possibility, we ran the fastest gait
from Glette et al. 10 times on ---L and measured an average speed of
only 12.95 cm/s $\pm0.93$ SD (vs. 17.8 cm/s measured on ROBIN) and a maximum of
13.8 cm/s (\tabref{results}). It thus appears that the ---L version of the robot is slower. Moreover, the HyperNEAT gait (14.5 cm/s) is faster than the best gait of Glette et al. on the ---L QuadraTot. Unfortunately, it was not possible to test the best HyperNEAT gait on the ROBIN QuadraTot. Because HyperNEAT outperformed the simple encoding from Glette et al. on the same robotic hardware, we tentatively
conclude that HyperNEAT produces faster gaits for the QuadraTot robot. This conclusion is supported by the fact that HyperNEAT also outperformed the encoding from Glette et al. in simulation.

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline
                                         &              & Simulated  & Real Vel. & Real Vel.  \\
                                         & Evaluations  & Velocity &    (---L)     & (ROBIN) \\
\hline
Parameterized gaits + optimization \cite{yos:clune}   &153    & --    & 5.8 & --\\
\hline
HyperNEAT in hardware \cite{yos:clune}                 & 180         & --         &   9.7  & --   \\
\hline
RL PoWER Spline \cite{haocheng}                         & 300         & --         &   11.1 & --\\
\hline
GA + simulator \cite{glette}             & 60000       & *16.7       &   13.8   & \textbf{17.8}  \\
\hline
HyperNEAT + simulator [this paper]                     & 40000       & \textbf{25.4}       &   \textbf{14.5} & --\\
\hline
\end{tabular}
\vspace{.35cm}
\caption{Velocities of evolved gaits in simulation and on two different copies of the QuadraTot robot. Subject to availability, data are reported from the experiments for this paper and three previous studies. Reported are the total number of evaluations per run, the average of the fastest gaits produced in each run in simulation, and the single fastest gait produced on the ---L and ROBIN copies of the QuadraTot robot (see text for their differences). *Instead of using the single fitness value reported in~\cite{glette}, we ran 19 additional runs and used the average fitness of those 20 runs. Velocities are in cm/s, and bold indicates the best performance. }  
\tablabel{results}
\end{center}
\end{table}

### 2012-11-07 paper_sean/robot.tex
THIS COMMIT CONTAINS DOUBLE-BLIND REVIEW CHANGES

We performed experiments on the QuadraTot quadrupedal robot platform~(\figref{quadratotWhiteBooties}-Left)~\cite{yos:clune}.
It has 9 degrees of freedom: two joints per leg and one joint that rotates along the robot's midline. %we def want a picture of the cool robot. I propose putting the picture of the robot on the left and the picture of the simulated robot on the right of the same figure. That also does a nice job of highlighting the theme of the paper (robot+sim) in one figure. Try to get this figure to show up as early as possible in the paper (e.g top of the 2nd page). 
The QuadraTot hardware designs and the software for this project are open source\footnote{A parts list, hardware CAD files, software (including the simulator), and gait videos are available at http://creativemachines.cornell.edu/evolved-quadruped-gaits}, and all hardware components are either off-the-shelf or 3D-printed. %These traits enable other labs to perform research using the same platform, which eases comparison of different gait-learning algorithms. 
There are results on the platform for nine different learning algorithms from three previous publications~\cite{yos:clune,glette,haocheng}.

>>>>

We performed experiments on the QuadraTot quadrupedal robot platform~(\figref{quadratotWhiteBooties}-Left)~\cite{yos:clune}.
It has 9 degrees of freedom: two joints per leg and one joint that rotates along the robot's midline. %we def want a picture of the cool robot. I propose putting the picture of the robot on the left and the picture of the simulated robot on the right of the same figure. That also does a nice job of highlighting the theme of the paper (robot+sim) in one figure. Try to get this figure to show up as early as possible in the paper (e.g top of the 2nd page). 
The QuadraTot hardware designs and the software for this project are open source\footnote{A parts list, hardware CAD files, software (including the simulator), and gait videos are available at (Link removed for double-blind review)}, and all hardware components are either off-the-shelf or 3D-printed. %These traits enable other labs to perform research using the same platform, which eases comparison of different gait-learning algorithms. 
There are results on the platform for nine different learning algorithms from three previous publications~\cite{yos:clune,glette,haocheng}.
